---
title: "scAtlas SubEndothelial cells subset"
date: "`r format(Sys.time(),  '%d %B %Y')`"
author: "Mohamed Hassan"
output: 
  html_notebook:
    theme: cerulean
    toc: true
    toc_depth: 3
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = T, results = "hide")
require("knitr")
setwd("/home/lucamannino/R_Projects/scAtlas/Analysis/SubECs/")
opts_knit$set(root.dir = "/home/lucamannino/R_Projects/scAtlas/Analysis/SubECs/")
```

```{r random-seed-setting}
set.seed(12345)
```

```{r loading-libraries, results='hide'}
library(tidyverse)
library(Seurat)
library(patchwork)
```


# 1. Reading the Vascular object
```{r}
Vascular_IntegratedObj <- readRDS("Vascular_IntegratedObj_annotated.RDS")
Vascular_IntegratedObj
```

# 2. Subsetting the SubECs cells
```{r}
Idents(Vascular_IntegratedObj) <- Vascular_IntegratedObj$RNA_snn_res.0.7

SubECs_Obj <- subset(Vascular_IntegratedObj, 
                      RNA_snn_res.0.7 %in% 
                        c("3", "7", "11", "12", 
                          "14", "18", "20", "22"))
SubECs_Obj
```


```{r}
UMAPPlot(SubECs_Obj, label = T)
```

```{r}
DefaultAssay(SubECs_Obj)<- "RNA"
```


## Preprocessing the Vascular object
```{r}
SubECs_Obj <- NormalizeData(SubECs_Obj, normalization.method = "LogNormalize", 
                             scale.factor = 10000)
SubECs_Obj <-  FindVariableFeatures(SubECs_Obj, selection.method = "vst", nfeatures = 2000)
SubECs_Obj <-  ScaleData(SubECs_Obj, vars.to.regress="nCount_RNA")
SubECs_Obj<-  RunPCA(SubECs_Obj, ndims.print = 1:50)
```


```{r}
ElbowPlot(SubECs_Obj, ndims = 50)
```


```{r}
### Determining the number of significant PCs
pct <- SubECs_Obj@reductions$pca@stdev / sum(SubECs_Obj@reductions$pca@stdev) * 100
cum <- cumsum(pct)
co1 <- which(cum > 90 & pct < 5)[1]
co2 <- sort(which((pct[1:length(pct) -1] - pct[2:length(pct)]) > 0.05), decreasing = T)[1] + 1
pcs = min(co1, co2)
sig_pcs <- 1:pcs
```


```{r}
SubECs_Obj <- FindNeighbors(SubECs_Obj, dims = sig_pcs)
```

```{r}
SubECs_Obj <- RunUMAP(SubECs_Obj, dims = sig_pcs)
UMAPPlot(SubECs_Obj, group.by = "Study")
UMAPPlot(SubECs_Obj, group.by = "technology")
```


```{r}
pdf("BatchEffect_SubECs_Obj.pdf")
UMAPPlot(SubECs_Obj, group.by = "Study")
UMAPPlot(SubECs_Obj, group.by = "technology")
dev.off()
```


```{r, eval=FALSE}
saveRDS(SubECs_Obj, "SubECs_Obj_Unintegrated.RDS")
```


```{r}
library(harmony)
SubECs_IntegratedObj <- RunHarmony(object = SubECs_Obj,
                                     group.by.vars=c("samples","Study_chemistry"),
                                     verbose = T, assay.use= "RNA",max.iter.harmony = 100)
### Determining the number of significant PCs
pct <- SubECs_IntegratedObj@reductions$harmony@stdev / sum(SubECs_IntegratedObj@reductions$harmony@stdev) * 100
cum <- cumsum(pct)
co1 <- which(cum > 90 & pct < 5)[1]
co2 <- sort(which((pct[1:length(pct) -1] - pct[2:length(pct)]) > 0.05), decreasing = T)[1] + 1
pcs = min(co1, co2)
harmony_sig_pcs <- 1:pcs



```{r}
SubECs_IntegratedObj <- RunUMAP(SubECs_IntegratedObj,
                                  dims = c(1:15),
                                  reduction = "harmony", verbose = T)
```




## Quality of the integration
```{r}
UMAPPlot(SubECs_IntegratedObj, group.by = "Study_chemistry") + NoLegend()
```


```{r}
UMAPPlot(SubECs_IntegratedObj, group.by = "Study")
UMAPPlot(SubECs_IntegratedObj, group.by = "Condition")
```

```{r}
SubECs_IntegratedObj <- FindNeighbors(SubECs_IntegratedObj,
                                        dims = c(1:15),
                                        reduction = "harmony", verbose = T)
```


```{r, fig.width=12}
res <- seq(0.1, 1, 0.1)
SubECs_IntegratedObj <- FindClusters(SubECs_IntegratedObj,resolution = res)
library(clustree)
clustree(SubECs_IntegratedObj)

```



```{r, fig.width=20, fig.height=10}
DimPlot(SubECs_IntegratedObj, 
         group.by = grep(pattern = "^RNA_snn", 
                         x = colnames(SubECs_IntegratedObj@meta.data)), 
         label = T, ncol = 5) & NoLegend()
```




```{r}
DefaultAssay(SubECs_IntegratedObj) <- "RNA"
Idents(SubECs_IntegratedObj) <- SubECs_IntegratedObj$RNA_snn_res.0.2

SubECs_markers <- FindAllMarkers(SubECs_IntegratedObj,
                              only.pos = T,
                              min.pct = 0.15, 
                              logfc.threshold = 0.3)
```


## Annotation
```{r}
new_names = c(
  '0'='FAP-like Endothelial cells', 
  '1'='Adipocyte-like Endothelial cells', 
  '2'='Activated Vascular Endothelial cells', 
  '3'='Myeloid-like Endothelial cells 1', 
  '4'='Lymphoid-like Endothelial cells',
  '5'='Mesenchymal-like Endothelial cells',
  '6'='Myeloid-like Endothelial cells 2',
  '7'='FAP-like Endothelial cells')

Idents(SubECs_IntegratedObj) <- SubECs_IntegratedObj$RNA_snn_res.0.2

SubECs_IntegratedObj$SubECs_labels <- recode(Idents(SubECs_IntegratedObj),
                                                 !!!new_names)
```

