---
title: "Emont_Integration - Emont's data"
date: "`r format(Sys.time(),  '%d %B %Y')`"
author: "Mohamed Hassan"
output: 
  html_notebook:
    theme: cerulean
    toc: true  
    toc_depth: 3
editor_options: 
  markdown: 
    wrap: 72
---


```{r}
library(Seurat)
library(tidyverse)
```

```{r}
Emont_all <- readRDS("/home/lucamannino/R_Projects/scAtlas/Mouse/Emont_only/mouse_all.rds")
Emont_all
```

## Removing unwanted assays
```{r}
Emont_all@assays$SCT <- NULL
Emont_all@assays$integrated <- NULL
```


```{r}
Emont_data <- subset(Emont_all, depot == "ING")
Emont_data
```

```{r}
Emont_data <- scCustomize::Add_Mito_Ribo_Seurat(Emont_data, 
                                                species = "mouse", 
                                                mito_name = "Mt", ribo_name = "Rb",
                                                mito_ribo_name = "Mt_Rb")
```

## Removing ribosomal genes
```{r}
Emont_data
Emont_data <- subset(Emont_data,
                     features = grep(pattern = "^Rp[l|s]", x = rownames(Emont_data),
                                       invert = T))
Emont_data
```


## Removing mitochondrial genes
```{r}
Emont_data
Emont_data <- subset(Emont_data,
                     features = grep(pattern = "^mt-", x = rownames(Emont_data),
                                       invert = T))
Emont_data
```

## Removing Malat1 and Neat1
```{r}
Emont_data
Emont_data <- subset(Emont_data,
                     features = grep(pattern = "Malat1", x = rownames(Emont_data), invert = T))
Emont_data <- subset(Emont_data,
                     features = grep(pattern = "Neat1", x = rownames(Emont_data), invert = T))

Emont_data
```

## Removing hemoglobin genes
```{r}
Emont_data
Emont_data <- subset(Emont_data,
                       features = grep(pattern = "^Hbb", x = rownames(Emont_data), invert = T))
Emont_data
```


## Scatterplot (UMIs vs Genes)
```{r}
FeatureScatter(Emont_data, feature1 = "nFeature_RNA", feature2 = "nCount_RNA", 
               group.by = "sample", raster = F)
```


```{r}
Emont_data
Emont_data <- subset(Emont_data,
                        subset = nFeature_RNA > 200 & nFeature_RNA < 9000 &
                         nCount_RNA > 200 & nCount_RNA < 50000)
Emont_data
```


```{r}
FeatureScatter(Emont_data, feature1 = "nFeature_RNA", feature2 = "nCount_RNA",
               group.by = "sample", raster = F)
```

```{r}
lin_mod <- lm(log(Emont_data@meta.data$nFeature_RNA) ~ log(Emont_data@meta.data$nCount_RNA))
sub <- lin_mod$model[lin_mod$residuals < -0.5,]
plot(log(Emont_data@meta.data$nCount_RNA),log(Emont_data@meta.data$nFeature_RNA), xlab ="log(nCount_RNA)",ylab="log(nFeature_RNA)",pch=20,col="grey") +
abline(lin_mod,col="red",lwd=3) +
points(sub[,2], sub[,1],pch=20,col="blue")
tokeep <- as.numeric(names(lin_mod$residuals[lin_mod$residuals >= -0.5]))
Emont_data <- subset(Emont_data, cells = tokeep)
```


```{r}
Emont_data
```


```{r}
Emont_data <- NormalizeData(Emont_data, normalization.method = "LogNormalize", scale.factor = 10000)
Emont_data <-  FindVariableFeatures(Emont_data, selection.method = "vst", nfeatures = 2000)

Emont_data <-  ScaleData(Emont_data, vars.to.regress=("nCount_RNA"))
Emont_data <-  RunPCA(Emont_data, features = VariableFeatures(object = Emont_data),
                        ndims.print = 1:50)


pct <- Emont_data@reductions$pca@stdev / sum(Emont_data@reductions$pca@stdev) * 100
cum <- cumsum(pct)
co1 <- which(cum > 90 & pct < 5)[1]
co2 <- sort(which((pct[1:length(pct) -1] - pct[2:length(pct)]) > 0.05), decreasing = T)[1] + 1
pcs = min(co1, co2)
sig_pcs <- 1:pcs
```


```{r}
Emont_data <- RunUMAP(Emont_data, dims = sig_pcs, reduction = "pca")
DimPlot(Emont_data, label = T, group.by = "sample", raster = F)
```


```{r, eval=FALSE}
pdf("Emont_BatchEffect.pdf")
UMAPPlot(Emont_data, label = F, group.by = "sample", raster = F)
dev.off()
```



```{r, eval=FALSE}
saveRDS(Emont_data, "Emont_dataObj_BeforeIntegration.RDS")
```


# Data Integration
```{r}
Emont_data$sample <- as.factor(Emont_data$sample) 
library(harmony)
Emont_Integrated <- RunHarmony(object = Emont_data,
                                        group.by.vars= "sample",
                                        assay.use= "RNA",max.iter.harmony = 100)

pct <- Emont_Integrated@reductions$harmony@stdev / sum(Emont_Integrated@reductions$harmony@stdev) * 100
cum <- cumsum(pct)
co1 <- which(cum > 90 & pct < 5)[1]
co2 <- sort(which((pct[1:length(pct) -1] - pct[2:length(pct)]) > 0.05), decreasing = T)[1] + 1
pcs = min(co1, co2)
harmony_sig_pcs <- 1:pcs
```


```{r}
Emont_Integrated <- RunUMAP(Emont_Integrated, 
                                     dims = c(1:30),
                                     reduction = "harmony", seed.use = 12345)

DimPlot(Emont_Integrated, split.by ="sample", group.by= "sample", label = T, ncol = 6, repel = F) & NoLegend()

DimPlot(Emont_Integrated, group.by= "sample")
```



```{r}
Emont_Integrated <- FindNeighbors(Emont_Integrated, dims = c(1:30), reduction = "harmony")
res <- seq(0.1, 1, 0.1)
Emont_Integrated <- FindClusters(Emont_Integrated, resolution = res)
```


```{r, fig.width=12}
library(clustree)
clustree(Emont_Integrated)
```

```{r, fig.width=20, fig.height=10}
UMAPPlot(Emont_Integrated, 
         group.by = grep(pattern = "^RNA_snn", 
                         x = colnames(Emont_Integrated@meta.data)), 
         label = T, ncol = 5, raster = F) & NoLegend()
```


```{r}
DefaultAssay(Emont_Integrated) <- "RNA"
Idents(Emont_Integrated) <- Emont_Integrated$RNA_snn_res.0.5

Emont_BigObj_Markers_res05 <- FindAllMarkers(Emont_Integrated,
                                only.pos = T,
                                min.pct = 0.15,
                                logfc.threshold = 0.3)

xlsx::write.xlsx(as.data.frame(Emont_BigObj_Markers_res05), "Emont_reIntegrated_factors_BigObj_Markers_res05.xlsx")
```


# Emont big object annotation
```{r}
Idents(Emont_Integrated) <- Emont_Integrated$RNA_snn_res.0.5

new_names <- c("0" = "Adipocytes", 
               "1" = "ASPCs", 
               "2" = "ASPCs", 
               "3" = "Macrophages and Monocytes", 
               "4" = "Adipocytes", 
               "5" = "Macrophages and Monocytes", 
               "6" = "Macrophages and Monocytes", 
               "7" = "Epithelial cells", 
               "8" = "ASPCs", 
               "9" = "Endothelial cells", 
               "10" = "Epithelial cells", 
               "11" = "Neutrophils",
               "12" = "Epithelial cells",
               "13" = "Macrophages and Monocytes",
               "14" = "Endothelial cells",
               "15" = "T and NK cells",
               "16" = "Adipocytes",
               "17" = "Mural cells",
               "18" = "Lymphatic Endothelial cells",
               "19" = "Epithelial cells")

Emont_Integrated$main_annotation <- recode(Idents(Emont_Integrated), 
                                           !!!new_names)
```

```{r}
DimPlot(Emont_Integrated, raster = F, group.by = "main_annotation", label = T, repel = T)
```


```{r}
saveRDS(Emont_Integrated, "Emont_Integrated_BigObj_annotated.RDS")
```


## Markers for the paper (Emont big object)
```{r}
DefaultAssay(Emont_Integrated) <- "RNA"
Idents(Emont_Integrated) <- Emont_Integrated$main_annotation

Emont_BigObj_final_markers <- FindAllMarkers(Emont_Integrated,
                                only.pos = T,
                                min.pct = 0.15,
                                logfc.threshold = 0.3)

xlsx::write.xlsx(as.data.frame(Emont_BigObj_final_markers),
                 "EmontBigObj_Markers_final.xlsx")
```

```{r}
Emont_Integrated$main_annotation <- factor(Emont_Integrated$main_annotation, 
                                           levels = c("ASPCs", "Adipocytes", 
                                                      "Endothelial cells", 
                                                      "Lymphatic Endothelial cells", 
                                                      "Mural cells", 
                                                      "Neutrophils", 
                                                      "Macrophages and Monocytes", 
                                                      "T and NK cells",
                                                      "Epithelial cells"))
```


```{r}
#### Reference from human big object

# c("ASPCs" = "steelblue4",
#   "Adipocytes" = "goldenrod4",
#   "Endothelial cells" = "red4",
#   "Lymphatic Endothelial cells" = "forestgreen",
#   "Mural cells" = "indianred4",
#   "Mast cells" = "darkslategrey",
#   "Macrophages and Monocytes" = "paleturquoise4",
#   "B cells" = "cadetblue",
#   "T and NK cells" = "violetred4")

mouse_bigobj_cluster_cols <- c("steelblue4",
                               "goldenrod4",
                               "red4",
                               "forestgreen",
                               "indianred4",
                               "cadetblue4",
                               "darkslategrey",
                               "violetred4", 
                               "midnightblue",
                               "steelblue")

DimPlot(Emont_Integrated, group.by = "main_annotation", cols = mouse_bigobj_cluster_cols)
```


```{r}
pdf("Emont_BigObj_color_coded_clusters_NoLegend.pdf", width = 10, height = 10)
DimPlot(Emont_Integrated, 
        group.by = "main_annotation", 
        cols = mouse_bigobj_cluster_cols, 
        pt.size = 0.1) + ggtitle("") + NoAxes() + NoLegend()
dev.off()
```



# Emont Vascular subset

## Vascular subset, includes LECs VSMCs Pericytes and Endothelial Cells
```{r}
Vascular_Emont <-  subset(Emont_Integrated, 
                          RNA_snn_res.0.5 %in% c("9", "14", "17", "18"))
```


```{r}
UMAPPlot(Vascular_Emont, label = T, raster = F)
```



## Vascular Subset Integration
```{r}
DefaultAssay(Vascular_Emont)<- "RNA"
set.seed(12345)
Vascular_Emont <- NormalizeData(Vascular_Emont, normalization.method = "LogNormalize", scale.factor = 10000)
Vascular_Emont <-  FindVariableFeatures(Vascular_Emont, selection.method = "vst", nfeatures = 2000)

Vascular_Emont <-  ScaleData(Vascular_Emont, vars.to.regress=("nCount_RNA"))
Vascular_Emont <-  RunPCA(Vascular_Emont, features = VariableFeatures(object = Vascular_Emont))


pct <- Vascular_Emont@reductions$pca@stdev / sum(Vascular_Emont@reductions$pca@stdev) * 100
cum <- cumsum(pct)
co1 <- which(cum > 90 & pct < 5)[1]
co2 <- sort(which((pct[1:length(pct) -1] - pct[2:length(pct)]) > 0.05), decreasing = T)[1] + 1
pcs = min(co1, co2)
vascular_sig_pcs <- 1:pcs

Vascular_Emont <- RunUMAP(Vascular_Emont, dims = vascular_sig_pcs)
```


```{r}
UMAPPlot(Vascular_Emont, label = T, raster = F, group.by = "sample")
```


```{r}
saveRDS(Vascular_Emont, "Vascular_Emont_Unitgerated.RDS")
```


## Vascular obj Integration
```{r}
Emont_Vascular_Integrated <- RunHarmony(object = Vascular_Emont,
                                        group.by.vars="sample",
                                        assay.use= "RNA",max.iter.harmony = 100)

pct <- Emont_Vascular_Integrated@reductions$harmony@stdev / sum(Emont_Vascular_Integrated@reductions$harmony@stdev) * 100
cum <- cumsum(pct)
co1 <- which(cum > 90 & pct < 5)[1]
co2 <- sort(which((pct[1:length(pct) -1] - pct[2:length(pct)]) > 0.05), decreasing = T)[1] + 1
pcs = min(co1, co2)
harmony_vascular_sig_pcs <- 1:pcs
```

```{r}
set.seed(12)
Emont_Vascular_Integrated <- RunUMAP(Emont_Vascular_Integrated, 
                                     dims = c(1:30),
                                     reduction = "harmony", seed.use = 12)

DimPlot(Emont_Vascular_Integrated, group.by= "sample", label = T) & NoLegend()
```

```{r}
Emont_Vascular_Integrated <- FindNeighbors(Emont_Vascular_Integrated, 
                               dims = c(1:30), 
                               reduction = "harmony")
res <- seq(0.1, 1, 0.1)
Emont_Vascular_Integrated <- FindClusters(Emont_Vascular_Integrated, resolution = res)
```

```{r, fig.width=9, fig.height=9}
library(clustree)
clustree(Emont_Vascular_Integrated)
```


```{r, fig.width=20, fig.height=10}
UMAPPlot(Emont_Vascular_Integrated, 
         group.by = grep(pattern = "^RNA_snn", 
                         x = colnames(Emont_Vascular_Integrated@meta.data)), 
         label = T, ncol = 5, raster = F) & NoLegend()
```


```{r}
Idents(Emont_Vascular_Integrated) <- Emont_Vascular_Integrated$RNA_snn_res.0.5

# Running marker genes
DefaultAssay(Emont_Vascular_Integrated) <-"RNA"

Emont_VascularObj_Markers_res05 <- FindAllMarkers(Emont_Vascular_Integrated,
                                only.pos = T,
                                min.pct = 0.15,
                                logfc.threshold = 0.3)

xlsx::write.xlsx(as.data.frame(Emont_VascularObj_Markers_res05),
                 "/home/lucamannino/R_Projects/scAtlas/Mouse/Emont_only/Emont_VascularObj_Markers_res05.xlsx")
```


# Emont big object annotation
```{r}
Idents(Emont_Vascular_Integrated) <- Emont_Vascular_Integrated$RNA_snn_res.0.5

new_names <- c("0" = "Capillary Endothelial cells 1", 
               "1" = "Arterial Endothelial cells", 
               "2" = "Arterial Endothelial cells", 
               "3" = "Pericytes", 
               "4" = "Venous Endothelial cells", 
               "5" = "Capillary Endothelial cells 2", 
               "6" = "Lymphatic Endothelial cells 1", 
               "7" = "Lymphatic Endothelial cells 2", 
               "8" = "Venous Endothelial cells", 
               "9" = "SubEndothelial cells", 
               "10" = "SubEndothelial cells", 
               "11" = "VSMCs",
               "12" = "Proliferating Endothelial cells", 
               "13" = "Arterial Endothelial cells", 
               "14" = "Arterial Endothelial cells")

Emont_Vascular_Integrated$vascular_labels <- recode(Idents(Emont_Vascular_Integrated), 
                                           !!!new_names)
```

```{r}
DimPlot(Emont_Vascular_Integrated, group.by = "vascular_labels", label = T, repel = T) + NoLegend()
```

```{r}
Emont_Vascular_Integrated$vascular_labels <- factor(Emont_Vascular_Integrated$vascular_labels,
                                                    levels = c("Venous Endothelial cells", 
                                                               "Capillary Endothelial cells 1", 
                                                               "Capillary Endothelial cells 2", 
                                                               "Arterial Endothelial cells", 
                                                               "SubEndothelial cells",
                                                               "Lymphatic Endothelial cells 1",
                                                               "Lymphatic Endothelial cells 2", 
                                                               "Pericytes", "VSMCs", 
                                                               "Proliferating Endothelial cells"))
```


```{r}
Idents(Emont_Vascular_Integrated) <- Emont_Vascular_Integrated$vascular_labels

# Running marker genes
DefaultAssay(Emont_Vascular_Integrated) <-"RNA"

Emont_VascularObj_Markers_final <- FindAllMarkers(Emont_Vascular_Integrated,
                                only.pos = T,
                                min.pct = 0.15,
                                logfc.threshold = 0.3)

xlsx::write.xlsx(as.data.frame(Emont_VascularObj_Markers_final), 
                "/home/lucamannino/R_Projects/scAtlas/Mouse/Emont_only/Emont_VascularObj_Markers_final.xlsx")

```


```{r}
vascular_mouse_colors <- c("navy",
                           "#eea9b8",
                           "#8b636c",
                           "#cd3333",
                           "#8b6508",
                           "forestgreen",
                           "darkgreen",
                           "#ffa54f",
                           "#cd853f",
                           "lightblue4")

pdf("Emont_mouse_Vascular_labels.pdf", width = 10, height = 10)
DimPlot(Emont_Vascular_Integrated, 
        group.by = "vascular_labels", 
        cols = vascular_mouse_colors,
        pt.size = 0.1) + NoAxes()
dev.off()
```










