---
title: "scAtlas SAT Integration and annotation"
date: "`r format(Sys.time(),  '%d %B %Y')`"
author: "Mohamed Hassan"
output: 
  html_notebook:
    theme: cerulean
    toc: true  
    toc_depth: 3
editor_options: 
  markdown: 
    wrap: 72
---

# 1. Reading the Big object


```{r}
All_Data_Atlas <- readRDS(file="../All_Data_Atlas.RDS")
All_Data_Atlas
```

## Adding a column combining the study and technology (Will be used for integration)
```{r}
All_Data_Atlas$Study_chemistry <- str_c(All_Data_Atlas$Study,
                                        All_Data_Atlas$Chemistry, sep = "_")
```

```{r}
DefaultAssay(All_Data_Atlas)<- "RNA"
```

## Adding Cell cycle score
```{r}
library("RCurl")
library(AnnotationHub)

cc_file <- getURL("https://raw.githubusercontent.com/hbc/tinyatlas/master/cell_cycle/Homo_sapiens.csv") 
cell_cycle_genes <- read.csv(text = cc_file)


# Connect to AnnotationHub
ah <- AnnotationHub()

# Access the Ensembl database for organism
ahDb <- query(ah, 
              pattern = c("Homo sapiens", "EnsDb"), 
              ignore.case = TRUE)

# Acquire the latest annotation files
id <- ahDb %>%
        mcols() %>%
        rownames() %>%
        tail(n = 1)

# Download the appropriate Ensembldb database
edb <- ah[[id]]

# Extract gene-level information from database
annotations <- genes(edb, 
                     return.type = "data.frame")

# Select annotations of interest
annotations <- annotations %>%
        dplyr::select(gene_id, gene_name, seq_name, gene_biotype, description)



# Get gene names for Ensembl IDs for each gene
cell_cycle_markers <- dplyr::left_join(cell_cycle_genes, annotations, by = c("geneID" = "gene_id"))

# Acquire the S phase genes
s_genes <- cell_cycle_markers %>%
        dplyr::filter(phase == "S") %>%
        pull("gene_name")
        
# Acquire the G2M phase genes        
g2m_genes <- cell_cycle_markers %>%
        dplyr::filter(phase == "G2/M") %>%
        pull("gene_name")

All_Data_Atlas <- CellCycleScoring(All_Data_Atlas, 
                                   s.features = s_genes, 
                                   g2m.features = g2m_genes)
```



## Quality check
```{r, fig.width=10, fig.height=10}
VlnPlot(All_Data_Atlas,features = c("nCount_RNA", "nFeature_RNA"), group.by ="Study_chemistry", pt.size = 0) + NoLegend()
```

```{r, fig.height=7, fig.width=12}
FeatureScatter(All_Data_Atlas, 
               feature1 = "nCount_RNA", feature2 = "nFeature_RNA", group.by = "Study")
```

```{r}
All_Data_Atlas_unfiltered <- All_Data_Atlas
```


All_Data_Atlas_unfiltered
All_Data_Atlas <- subset(All_Data_Atlas, 
                         nCount_RNA < 60000 & nFeature_RNA < 10000)
All_Data_Atlas
```

# 2. Data Processing
```{r}
DefaultAssay(All_Data_Atlas)<- "RNA"
All_Data_Atlas <- NormalizeData(All_Data_Atlas, normalization.method = "LogNormalize", 
                                scale.factor = 10000)
All_Data_Atlas <-  FindVariableFeatures(All_Data_Atlas, selection.method = "vst", 
                                        nfeatures = 2000)
All_Data_Atlas <-  ScaleData(All_Data_Atlas, vars.to.regress =  "nCount_RNA")
```


```{r}
All_Data_Atlas <-  RunPCA(All_Data_Atlas, features = VariableFeatures(object = All_Data_Atlas), ndims.print = 1:40)
```


```{r}
ElbowPlot(All_Data_Atlas, ndims = 50)

### Determining the number of significant PCs
pct <- All_Data_Atlas@reductions$pca@stdev / sum(All_Data_Atlas@reductions$pca@stdev) * 100
cum <- cumsum(pct)
co1 <- which(cum > 90 & pct < 5)[1]
co2 <- sort(which((pct[1:length(pct) -1] - pct[2:length(pct)]) > 0.05), decreasing = T)[1] + 1
pcs = min(co1, co2)
sig_pcs <- 1:pcs
```


```{r}
All_Data_Atlas <- FindNeighbors(All_Data_Atlas, 
                                          reduction = "pca", dims = sig_pcs)


All_Data_Atlas <- RunUMAP(All_Data_Atlas, dims = sig_pcs)

UMAPPlot(All_Data_Atlas, label = T, 
         group.by = c("Study", "technology", "orig.ident",
                      "Chemistry", "Study_chemistry", "Condition", "Phase"), 
         raster = F) + NoLegend()
```


# 3. Integration (Removing batch effect)
```{r}
All_Data_Atlas$orig.ident <- as.factor(All_Data_Atlas$orig.ident)

## Adding samples column for integration
All_Data_Atlas$samples<- as.character(All_Data_Atlas$orig.ident)

# Running the integration with Harmony
Atlas_IntegratedObj <- RunHarmony(All_Data_Atlas, 
                             group.by.vars= c("samples", "Study_chemistry"), 
                             verbose = T, assay.use= "RNA", max.iter.harmony = 100)

ElbowPlot(Atlas_IntegratedObj, reduction =  "pca", ndims = 50) +  ElbowPlot(Atlas_IntegratedObj, reduction =  "harmony", ndims = 50)


### Determining the number of significant PCs
pct <- Atlas_IntegratedObj@reductions$harmony@stdev / sum(Atlas_IntegratedObj@reductions$harmony@stdev) * 100
cum <- cumsum(pct)
co1 <- which(cum > 90 & pct < 5)[1]
co2 <- sort(which((pct[1:length(pct) -1] - pct[2:length(pct)]) > 0.05), decreasing = T)[1] + 1
pcs = min(co1, co2)
harmony_sig_pcs <- 1:pcs
```


```{r}
Atlas_IntegratedObj <- RunUMAP(Atlas_IntegratedObj, dims = harmony_sig_pcs,reduction = "harmony", verbose = T)
```


## Checking the quality of integration
```{r}
UMAPPlot(Atlas_IntegratedObj, group.by = "Study", raster= F)
UMAPPlot(Atlas_IntegratedObj, group.by = "technology", raster= F)
UMAPPlot(Atlas_IntegratedObj, group.by = "Study_chemistry", raster= F)
UMAPPlot(Atlas_IntegratedObj, group.by = "orig.ident", raster= F) + NoLegend()
```


```{r}
# UMAP split by study
DimPlot(Atlas_IntegratedObj, split.by ="Study",group.by= "orig.ident", raster = F,
       label = T, repel = F, ncol = 4) & NoLegend()
```



```{r}
Atlas_IntegratedObj<- FindNeighbors(Atlas_IntegratedObj, dims = harmony_sig_pcs, 
                                    reduction = "harmony", verbose = FALSE)
res <- seq(0.1, 3, by=0.1)
Atlas_IntegratedObj <- FindClusters(Atlas_IntegratedObj, verbose = FALSE, resolution = res)
```


## Selecting optimal resolution based on clustree
```{r, eval=FALSE}
library(clustree)
Clustree_Plot <- clustree::clustree(Atlas_IntegratedObj)
Clustree_Plot
```


```{r}
DimPlot(Atlas_IntegratedObj,  split.by = "Condition", ncol = 3 ,raster =F,label=T)
```

```{r}
DimPlot(Atlas_IntegratedObj,  group.by = "Condition",raster =F)
```



# Cluster Annotation
```{r}
Idents(Atlas_IntegratedObj) <- Atlas_IntegratedObj$RNA_snn_res.0.1

All_Atlas_markers <- FindAllMarkers(Atlas_IntegratedObj, 
                              only.pos = T, 
                              min.pct = 0.15, 
                              logfc.threshold = 0.3)


All_Atlas_sig_markers <- All_Atlas_markers %>% dplyr::filter(p_val_adj < 0.05)


Atlas_IntegratedObj_test_high_res <- Atlas_IntegratedObj
```


## Subclustering cluster 5 at resolution 0.1
```{r}
subclustering5 <- FindSubCluster(object = Atlas_IntegratedObj_test_high_res, 
                                 cluster = "5", 
                                 graph.name = "RNA_snn", subcluster.name = "Subcluster5",
                                 resolution = 1, algorithm = 1)
```



## Running marker genes
```{r}
DefaultAssay(subclustering5) <- "RNA"
Idents(subclustering5) <- subclustering5$Subcluster5

Atlas_markers <- FindAllMarkers(subclustering5,
                              only.pos = T,
                              min.pct = 0.15, 
                              logfc.threshold = 0.3)

top50 <- Atlas_markers %>% group_by(cluster) %>% slice_max(n = 50, order_by = avg_log2FC)
top100 <- Atlas_markers %>% group_by(cluster) %>% slice_max(n = 100, order_by = avg_log2FC)

library(writexl)
write_xlsx(as.data.frame(Atlas_markers),
                    "markers/markers_Atlas_WRST_15PCT_Log03_recovered_res01_cluster5_subclustered.xlsx")
```


## Final annotation
```{r}
new_names = c(
  '0'='ASPCs', 
  '1'='Macrophages and Monocytes', 
  '2'='T and NK cells', 
  '3'='Endothelial cells',
  '4'='Adipocytes', 
  '5_0'='Endothelial cells', '5_1'='Endothelial cells',
  '5_2'='Endothelial cells', '5_3'='Endothelial cells', 
  '5_4'='Endothelial cells', '5_5'='Endothelial cells', 
  '5_6'='Endothelial cells', '5_7'='Endothelial cells', 
  '5_8'='Endothelial cells', '5_9'='Endothelial cells', 
  '5_10'='ASPCs', '5_11'='Endothelial cells', 
  '5_12'='Endothelial cells', '5_13'='Endothelial cells', 
  '5_14'='ASPCs', '5_15'='Endothelial cells', 
  '5_16'='Endothelial cells', '5_17'='Macrophages and Monocytes', 
  '5_18'='Adipocytes', '5_19'='Endothelial cells', 
  '5_20'='Endothelial cells', '5_21'='Endothelial cells', 
  '6'='Mural cells', 
  '7'='Macrophages and Monocytes', 
  '8'='Mast cells', 
  '9'='B cells', 
  '10'='Adipocytes', 
  '11'='Lymphatic Endothelial cells')

Idents(Atlas_IntegratedObj) <- Atlas_IntegratedObj$Subcluster5

Atlas_IntegratedObj$final_annotation <- recode(Idents(Atlas_IntegratedObj),
                                                 !!!new_names)
```


```{r}
DimPlot(Atlas_IntegratedObj, group.by = "final_annotation", label = T, raster = F)
```





## Running marker genes
```{r}
DefaultAssay(Atlas_IntegratedObj) <- "RNA"
Idents(Atlas_IntegratedObj) <- Atlas_IntegratedObj$final_annotation

Atlas_markers <- FindAllMarkers(Atlas_IntegratedObj,
                              only.pos = T,
                              min.pct = 0.15, 
                              logfc.threshold = 0.3)

top50 <- Atlas_markers %>% group_by(cluster) %>% slice_max(n = 50, order_by = avg_log2FC)
top100 <- Atlas_markers %>% group_by(cluster) %>% slice_max(n = 100, order_by = avg_log2FC)

library(writexl)
write_xlsx(as.data.frame(Atlas_markers),
                    "markers/markers_Atlas_WRST_15PCT_Log03_res01_cluster5_subclustered_annotated_final4paper.xlsx")
```



```{r}
genes <- ggplot(Atlas_IntegratedObj@meta.data, aes(x=final_annotation, y=nFeature_RNA,fill=final_annotation)) + 
  geom_violin(trim=FALSE,scale="width") +
  geom_boxplot(width=0.1, fill="white",outlier.shape = NA) +
  labs(x="", y = "Genes") + theme_classic() + 
  theme (legend.position="none") + scale_fill_manual(values = bigobj_cluster_cols)+ theme(axis.text = element_text(size=15)) + ylab("")

genes
ggsave("Plots/nGenes_final_annotation_updated_colors.pdf", device="pdf",units="cm",width=52, height=15)
```

```{r}
UMIs <- ggplot(Atlas_IntegratedObj@meta.data, aes(x=final_annotation, y=nCount_RNA,fill=final_annotation)) +
  geom_violin(trim=FALSE,scale="width") +
  geom_boxplot(width=0.1, fill="white",outlier.shape = NA)+
  labs(x="", y = "UMIs") + theme_classic() + theme (legend.position="none") + scale_fill_manual(values = bigobj_cluster_cols)+ theme(axis.text = element_text(size=15)) + ylab("")# + scale_y_log10()
UMIs
ggsave("Plots/nUMIs_final_annotation_updated_colors.pdf", device="pdf",units="cm",width=52, height=15)
```
